---
title: "MIDTERM EXAM"
author: "Nicholas McInnis - 11290398"
output:
  pdf_document: default
  html_document: default
---

---

**Welcome to this midterm exam!**

Before doing anything, replace the author information in the YAML with your own information.

You can use the PDF version on ZoneCours for your comfort, but the answers must be provided in the Rmd file.

> You can start right now and you have 3 hours. You can use all the material made available for this course. Do not miss the deadline on ZoneCours, otherwise you will not be able to upload it, I cannot accept your files by emails or any other means. So, please give yourself enough time to upload your file on ZoneCours to avoid any potential issue (connections, etc.).

Please follow the directions for each question. **Only this Rmd file will be graded** (not Word, not the html version, etc.).

Once completed:

- Step 1. save this Rmd file in your session as an Rmd file by just clicking on the little disk icon in RStudio.cloud,

- Step 2. select it on the bottom-right panel by checking the little white box and then export it by clicking on the "more" button in the bottom-right panel,

- Step 3. and then upload it to ZoneCours.

# Important recommendations

- This midterm is graded over 25 points and is organized in three exercises.

- Have a look at the whole midterm first to gauge how much time you can devote to each question. This midterm values different skill sets. If you do not know the mathematical explanation, use plain English. I give partial credits.

- VERY IMPORTANT: do not focus only on the code, do not provide code without explanations of why you use this code. So, do not fall into the trap of just providing a code chunk thinking that it is the answer. It is only a partial answer.

- Individual exam, no communications allowed. If you are not sure about your understanding of a question, make assumptions.

- You have access to (only) all our class material.






\newpage 

---

# Exercise 1: Foundations Questions (Total points: 5)

Directions: no copying and pasting from your class material. Replace the mock text, which is there to give you a sense of how long/concise your answer should be.

## Question 1 (1 point)

**In your own words, can you explain the survivorship bias?**

Survivorship bias is a phenomenon where mathematical models are biased because they are making assumptions about the population using incomplete sample data. More specifically, it is only possible to collect data from "survivors" and therefore, statistical inferences from "survivors should not provide an accurate generalization of a true population that is comprised of both survivors and non-survisors, the latter of which we cannot collect data from. A good example is the plane and bullet holes example, where data on planes can only be collected from returning planes, and we will never know exactly where the bullter holes are for the non-surviving planes. Intuitively, we can argue that planes shot down received bullter holes in critical areas, such as the cockpit and engines, however we cannot support this statictically because of the survivorship bias.



## Question 2 (1 point)

**In your own words, can you explain the Central Limit Theorem?**

In my words, the central limit theorem describes how as we increase the sample size, the mean of the sample will converge towards the population (normal distribution) and will be more evenely distributed. This phenomenon can also be interpreted as the law of big numbers, and it the phenomenon can be observed using a histogram. In order for the phenomenom to occur, sampling observations must be independent from one another, be without replacement, and be selected using random sampling techniques.



## Question 3 (1 point)

**In your own words, explain TSS and RSS, why are they useful?**

TSS (total sum of squares) measures the difference between individual observations of the dependent variable (y) and their sample mean (y-bar), which represents the total variance in y. RSS on the other hand, measures the difference between individual observations of the dependent variable (y) and their corresponding fitted observation of the dependent by the model (y-hat). In other words, RSS represents the model error.

Together, RSS and TSS can be useful for finding F, which is useful for hypothesis testing and assessing model significance. They are also useful for finding R-squared, using the formula : R-squared = 1-(RSS/TSS) , which is key for assessing model significance. RSS can be divided by degrees of freedom to find the residual standard errors (RSE = sqrt(RRS/(n-2)), an effective metric for assessing the accuracy of the model.



## Question 4 (1 point)

**In your own words, explain what a confidence interval is.**

My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. 

## Question 5 (1 point)

**In your own words, explain hypothesis testing.**

My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. 







\newpage

---

# Exercise 2 (Total points: 8)

Let us imagine we are in 1999, and you have been selected as a referee for this submitted article to our academic journal: "Effects of age at entry, knowledge intensity, and imitability on international growth" Autio, Erkko; Sapienza, Harry J; Almeida, James G., submitted to Academy of Management Journal; Vol. 43, Iss. 5, 909-924.

Here is the link to the full version of the article, copy and paste it in your browser: <https://www-proquest-com.proxy2.hec.ca/docview/199843639?accountid=11357>

In the space below, **analyze this article with your new analytical skills (you may want to focus on the methods and results sections, pp. 914-918).**

In particular, highlight the contributions, the research question, the results, the models accuracy and the validity of the empirical estimations. For instance, do you suspect multicollinearity? Is there anything missing, something you would add to be sure? Use all the steps you have learned in the first 5 sessions of our course to assess the information you have in this article, the information you do not have, and maybe the steps that the authors did not do but would have been great to do.

Would you recommend:

- accept
- revise and resubmit with minor revisions
- revise and resubmit with major revisions
- reject

> Type your analysis below. Replace the mock text.

My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. 

My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. 

My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. 

My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. My response here. 

> End of your answer






\newpage

---

# Exercise 3 (Total points: 12)


Costel Calin and Kevin Buterbaugh have published an article titled "Male  versus Female Career Ambassadors: Is the US Foreign Service Still Biased?" in *Foreign Policy* (Volume 15, issue 2, April 2019). Here is the abstract of the original study:

> "This paper examines the appointment of male versus female career diplomats to ambassadorial posts. We assess the role played by ambassadors’ individual characteristics, including education, marital status, and number of children, and host countries’ characteristics, such as quality of life and regime type, in determining if a male or female is appointed to ambassadorial positions to represent the United States in foreign countries. The time frame of this study is the entire presidencies of Bill Clinton and George W. Bush (1993–2008), during which 603 career diplomats were appointed as ambassadors. The study provides empirical evidence that there remain significant differences between women and men serving as ambassadors. Female ambassadors are more likely to be single and have no children and are less likely to be Ivy League graduates than male ambassadors. Furthermore, they are more likely to be appointed to countries with lower quality of life and better human rights records. Finally, time plays a role in the likelihood of a woman being appointed as ambassador."

Just to give you an idea about their models:

|        Variable       |     Model 1    |     Model 2     |     Model 3     |
|:---------------------:|:--------------:|:---------------:|:---------------:|
|     Marital status    | 1.991*** (.41) | 2.061*** (.408) | 2.046*** (.408) |
|        Children       |  .718** (.360) |  .705** (.359)  |  .729** (.361)  |
|  Education Ivy League |   .397 (.318)  |        -        |        -        |
|       Graduate 1      |        -       |   0.74 (.200)   |        -        |
|       Graduate 2      |        -       |        -        |   -.090 (.302)  |
|          Time         |  -.062* (.032) |  -.060* (.032)  |  -.059* (.032)  |
|  Quality of life HDI  | 1.710** (.861) |  1.813** (.861) |  1.805** (.860) |
|       Inf.mort.       |        -       |        -        |        -        |
|  GDP/capita (logged)  |        -       |        -        |        -        |
| Regime type Emp.Index | -.104** (.043) |  -.104** (.043) |  -.104** (.043) |
|         Polity        |        -       |        -        |        -        |
| Power of host country |   .094 (.091)  |   .093 (.090)   |   .090 (.090)   |
|  Secr. of state (sex) |   .261 (.294)  |   .294 (.293)   |   .297 (.294)   |
|           N           |       455      |       456       |       456       |
|      $\chi^2$         |      96.68     |      95.37      |      95.32      |
|          P <          |      0.000     |      0.000      |      0.000      |
|         $R^2$         |      .199      |       .196      |       .196      |



Both authors have heard about your great skills as a data analyst. They have updated their dataset with a new variable called "bonus". It captures the yearly bonuses ambassadors receive to compensate for some of the difficulties for living in their host country (cost of living, potential threats, etc.). Apparently, this calculation is done by AI. Both authors wonder whether they could use their updated dataset to do some new analyses. They are worried about biases in the bonus calculation.

Here is a list of the variables:

| Variable name | Definition
|---------------|----------------------------------------------------------------------|
| sexofamb      | ambassador's sex, 1: female, 0: male
| married       | ambassador's marital status
| children      | 1; if the ambassador has 1 or more children, 0: otherwise
| HDI           | human development index of host country
| infmort       | infant mortality in host country
| ivy           | Ivy league school graduate
| time          | number of years in the diplomacy
| sexofsecr     | Secretary of State's sex, 1: female, 0: male
| loggdp2       | Log GDP of host country
| graduate1     | Ambassador's education: 0 = undergraduate, 1 = Masters, 2 = PhD
| graduate2     | Ambassador's education: 0 = undergraduate, 1 = graduate

They put their dataset on a highly secured server, which you can access through this code (run the code chunk by clicking on the little green triangle on the top-right of the code chunk):

```{r, message = FALSE, warning = FALSE}
if (!require("dplyr")) install.packages("dplyr")
library(dplyr)
library(readr)
mydata1 <- readr::read_csv("https://warin.ca/datalake/courses_data/qmibr/09128734/us_confidential_2.csv")

# uncomment (=remove the hashtag) to view your dataframe
#View(mydata1)
```

## Question 1 (1 point)

The new model (let's call it "m4") that Costel Calin and Kevin Buterbaugh want you to test is the following simple linear regression: **"Can we explain the dependent variable "bonus" with the risk exposure for the ambassadors , the latest being measured by the Human Development Index (HDI)?"** Write the code below and explain your results in terms of model accuracy.

```{r, eval = FALSE}
m4 <- 
  
  
```


Explanations:




## Question 2 (2 points)

**Now, you tell your co-authors that it is also - and maybe more importantly - about model validity. Define the three important concepts you would like to analyze in order to assess the model's validity. Then, run the respective tests in R in the following code chunks. Explain the results.**

Concept 1: Define the concept and write the code below. Explain your results.

```{r, eval = FALSE}
m4 <- 
  
  
```


Concept 2: Define the concept and write the code below. Explain your results.

```{r, eval = FALSE}
m4 <- 
  
  
```



Concept 3: Define the concept and write the code below. Explain your results.

```{r, eval = FALSE}
m4 <- 
  
  
```


## Question 3 (1 point)

**To fix for some issues, you propose to look into a new model (let's call it "m5") considering HDI and graduate2 as your independent variables. Write the code below and explain your results in terms of model accuracy.**

```{r, eval = FALSE}
m5 <- 
  
  
  
```


## Question 4 (1.5 point)

**Compare m4 and m5, which one is better? Explain your options and results.**

```{r, eval = FALSE}



```



## Question 5 (2 points)

**Bonuses have also been used as a policy to make the ambassador's career more attractive to women. Are women really receiving higher bonuses? Hint: look at two independent variables: the male / female dummy ("sexofamb" variable) while considering the ambassadors' experience ("time" variable). Write the code below and explain your results in terms of model accuracy.**


```{r, eval = FALSE}
m6 <- 
  
  
  
```

**Write below the equations of the estimated bonus for men and separately for women.**



**What can you conclude?**


## Question 6 (2 points)

**You propose to do another interaction-based model, this time between sexofamb and sexofsecr. Write the code below and explain your results in terms of model accuracy.**

```{r, eval = FALSE}
m7 <- 
  
  
```

**Write below the equations of the possible scenarios (assume all the p-values are good at at least 95% for simplification purposes). Does a female Secretary of State gives higher bonuses to male or female ambassadors? Does a male secretary of State gives higher bonuses to a male or a female ambassador?**


## Question 7 (2.5 points)

**What would be your next steps for instance in terms of models, code, assessment?**






