---
title: "mock_midterm_exam"
author: "Thierry Warin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to this midterm exam.

You can start right now and you have 3 hours. Do not miss the deadline on ZoneCours, otherwise you will not be able to upload it, I cannot accept your files by emails or any other means. So, please give yourself enough time to upload your file on ZoneCours to avoid any potential issue (connections, etc.).

Please follow the directions for each questions. Only this Rmd file will be graded (not Word, not the html version, etc.).

Once completed:

Step 1. save this Rmd file in your session as an Rmd file by just clicking on the little disk pictogram in RStudio.cloud,

Step 2. select it on the bottom-right panel by checking the little white box and then export it by clicking on the "more" button in the bottom-right panel,

Step 3. and then upload it to ZoneCours.

This midterm is organized in two parts: foundations and applications. Have a look at the whole midterm first to gauge how much time you can devote to each question, and then go for it! Applications will require more time.

VERY IMPORTANT: do not focus on the code, do not provide code without explanations of why you use this code. Your answers should be literal answers, and then you can provide the results based on your analysis in plain English. So, do not fall into the trap of just providing a code chunk thinking that it is the answer. It is only a partial answer. I want to know your thinking, your analysis, and then I want to see if you know the R command that would provide provide the answer based on your analysis in plain English.

# 1. Foundations Questions

## Question 1

Can you fix the following code?

Note: I only want numbers with two decimals to make a nice table.

```{r}
poverty <- readr::read_tsv("https://warin.ca/datalake/courses_data/qmibr/session4/poverty.txt")

model1 <- lm(poverty$Poverty ~ poverty$Graduates)

stargazer(model1, type="text", header=FALSE)
```


## Question 2

For further examples, see the respective sessions' applications.

# 2. Applications Questions

## question 1

You have been selected as a referee for this submitted article to our academic journal: "Internationalization and performance of small- and medium-sized enterprises."

In a maximum of 500 words, analyze this article . Highlight the contributions and the validity of the empirical estimation. Use all the steps you have learned in the first 5 sessions of our course. Would you recommend:

- accept
- revise and resubmit with minor revisions
- revise and resubmit with major revisions
- reject

## Question 2

Your co-author has collected great data. Here is the location of the dataset: "https://www.warin.ca/datalake/mock_data.csv". You are the data scientist of the group and you are expected to make a great statistical contribution.

### Question 2.1

Draw an exploratory data analysis. DO NOT ONLY provide the code. Explain your goal, why it matters, explain the steps, and then do it in R in an R code chunk.

### Question 2.2

Your first research question is...

Build a model that corresponds to this RQ, analyze it in terms of accuracy and validity.

### Question 2.2




